---
title: "p8105_hw3_ay2543"
output: github_document
date: "2022-10-13"
---

Loading libraries:
```{r, message = FALSE}
library(tidyverse)
```

# Problem 1
```{r, message = FALSE}
library(p8105.datasets)
data("instacart")
```

## 1a
This is a preview of the dataset:
```{r}
head(instacart) %>% 
  knitr::kable()
```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations. Then, do or answer the following (commenting on the results of each):



* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

--
The instacart dataset has **`r nrow(instacart)` observations and `r ncol(instacart)` variables**.
Some key variables are *user_id, product_id, order_number, order_hour_of_day, order_dow*. 

## 1b
* How many aisles are there, and which aisles are the most items ordered from?

```{r}
aisle_count = instacart %>% 
  group_by(aisle) %>% 
  summarise(n_order = length(order_id)) %>% 
  arrange(desc(n_order))

aisle_count
```

There are `r length(unique(instacart$aisle))` aisles. Most items are ordered from the fresh vegetables aisle.

## 1c
* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
aisle_count %>%
  filter(n_order>10000) %>% 
  ggplot(aes(x = fct_reorder(aisle, n_order), y = n_order)) +
  geom_col() +
  coord_flip() +
  theme_classic() +
  labs(
    title = "Number of items ordered in each aisle",
    y = "Number of items ordered",
    x = "Aisle",
    caption = "Limited to aisles with > 10000 items ordered"
  )
```

## 1d

* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle) %>% 
  summarise(item = product_name)
```









